---
title: "Condition_Visualization"
author: "Rachele Spadafore"
date: "5/15/2021"
output: html_document
---
Markdown code for the visualization of the coral conditions (Mortality, Partial Mortality, Burial, Partial Burial, Bleaching, and Disease) over the observation period. The cumulative timeline plot includes only the 4 main conditions (Mortality, Burial, Bleaching and Disease)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Packages
```{r}
library(tidyverse)
library(janitor)
library(ggrepel)
library(ggthemes)
library(MASS)
library(merTools)
library(RColorBrewer)
library(ggplot2)
library(ggthemes)
library(ranger)
library(dplyr)
library(ggfortify)
library(ggsci)
library(ggpubr)
```

Figure 2: Cumulative Line Plot Set Up
```{r}
#Load data
cum.line=read.csv("line.cum.csv")

#Level Months
cum.line$Month=factor(cum.line$Month, levels = c("Oct_13", "Nov_13", "Dec_13", "Jan_14", "Feb_14", "Mar_14", "Apr_14", "May_14", "Jun_14", "Jul_14", "Aug_14", "Sep_14", "Oct_14", "Nov_14", "Dec_14", "Jan_15", "Feb_15", "Mar_15","Jun_15", "Jul_15"))

#Transform to long format
cum.line.long.all=pivot_longer(cum.line, cols = 2:5, names_to= "Condition", values_to= "Individuals")

#Re name Conditions for consisitancy
cum.line.long.all$Condition<-recode_factor(cum.line.long.all$Condition, "Buried"="Burial")
cum.line.long.all$Condition<-recode_factor(cum.line.long.all$Condition, "Bleached"="Bleaching")

#Level Conditions
cum.line.long.all$Condition <- factor(cum.line.long.all$Condition, levels = c("Mortality", "Burial", "Bleaching", "Disease"))

```

Figure 2: Line Plot (Cumulative)
```{r}
#Set up plot
cum.timeline.plot=cum.line.long.all%>%
  ggplot(aes(x=Month, y=Individuals, group=Condition, color=Condition))+
  geom_line()+
  ylab('Number of Corals') +
  xlab('Month')+
  #ggtitle("Coral Conditions During Dredge Observation")+
  scale_color_npg()+
  geom_point(size=4)+
  geom_text(aes(label=Individuals), color="white", size=2.5)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))+
 
   #Dredge Line and Label
  annotate("segment", x = 2, xend = 18, y = 145, yend = 145, colour = "gray", size=1.5,
           #arrow = arrow(ends = "both", angle = 90, length = unit(.2,"cm"))
           )+
   annotate(geom = "text", x =10, y = 152, label = "Dredging", color = "black", size=3) +
  
  #Bleach Line and Label
  annotate("segment", x = 10, xend = 12, y = 95, yend = 95, colour = "gray",size=1.5, 
           #arrow = arrow(ends = "both", angle = 90, length = unit(.2,"cm"))
           )+
  annotate(geom = "text", x =11, y = 105, label = "Hyperthermal Sea Surface Temperatures", color = "black", size= 3)+
  
  #Disease Line and Label
  annotate("segment", x = 12, xend = 20, y = 120, yend = 120, colour = "gray", size=1.5,
           #arrow = arrow(ends = "both", angle = 90, length = unit(.2,"cm"))
           )+
  annotate(geom = "text", x =16, y = 130, label = "SCTLD Outbreak", color = "black", size =3)
  

cum.timeline.plot
```

Save Figure 2 as Jpg
```{r}
ggsave(filename="Figure_2.jpg", plot=cum.timeline.plot, device="jpg", width = 7, height = 5)

```


First Occurance Line Plot (Non-cumulative) 
```{r}
#Load data
first.line=read.csv("line.plot.csv")

#Level Months
first.line$Month=factor(first.line$Month, levels = c("Oct_13", "Nov_13", "Dec_13", "Jan_14", "Feb_14", "Mar_14", "Apr_14", "May_14", "Jun_14", "Jul_14", "Aug_14", "Sep_14", "Oct_14", "Nov_14", "Dec_14", "Jan_15", "Feb_15", "Mar_15","Apr_15","May_15","Jun_15", "Jul_15"))

#Transform to long format
first.line.long.all=pivot_longer(first.line, cols = 2:7 , names_to= "Condition", values_to= "Individuals")

#Level Conditions
first.line.long.all$Condition <- factor(first.line.long.all$Condition, levels = c("Mortality", "Buried", "Bleached", "Disease", "Partial.Mortality", "Partially.Buried"))

```

First Occurance Line Plot (Non-Cumulative)
```{r}
#Set up plot
first.timeline.plot=first.line.long.all%>%
  ggplot(aes(x=Month, y=Individuals, group=Condition, color=Condition))+
  geom_line()+
  ylab('Number of Corals') +
  xlab('Month')+
  ggtitle("Coral Conditions During Dredge Observation")+
  scale_color_npg()+
  #geom_point()+
  #geom_text(label=Individuals)
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))+
 
   #Dredge Line and Label
  annotate("segment", x = 1, xend = 22, y = 150, yend = 150, colour = "gray", arrow = arrow(ends = "both", angle = 90, length = unit(.2,"cm")))+
   annotate(geom = "text", x =11, y = 157, label = "Dredging", color = "black", size=3) +
  
  #Bleach Line and Label
  annotate("segment", x = 10, xend = 11, y = 100, yend = 100, colour = "gray", arrow = arrow(ends = "both", angle = 90, length = unit(.2,"cm")))+
  annotate(geom = "text", x =10.5, y = 110, label = "Bleaching Event", color = "black", size= 3)+
  
  #Disease Line and Label
  annotate("segment", x = 12, xend = 22, y = 125, yend = 125, colour = "gray", arrow = arrow(ends = "both", angle = 90, length = unit(.2,"cm")))+
  annotate(geom = "text", x =17, y = 135, label = "SCTLD Outbreak", color = "black", size =3)
  

first.timeline.plot
```

Save First_Timeline_plot as a jpg
```{r}
ggsave(filename="Timeline_Plot_Noncumulative.jpg", plot=first.timeline.plot, device="jpg", width = 7, height = 5)
```










