---
title: "Survival Analyses"
output: github_document
always_allow_html: true
---

Markdown code for the survival analyses of the Port of Miami data examining the main effects of channel (i.e. dredging impact), reef habitat, and direction. Survival analyses used to tease apart species specific differences as well as the impact of state (or condition) specifically whether the corals were observed to be buried, bleached or diseased. Each variable is coded as 0 (absence) or 1 (presence) as well as in combination in the states column.

```{r setup}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE, error = FALSE, fig.path='Figures/', dev=c('png', 'pdf'))
```

All figures exported to Figures/folder as png and pdf

```{r libraries}
library(knitr)
library(survival)
library(coda)
library(survminer)
library(dplyr)
library(ggplot2)
library(car)
library(tidyverse)
library(kableExtra)
#library(gtsummary) #not loading properly
```

```{r import data}
df_surv=read.csv("fl_surv_minusMIS.csv") #time by months; excludes missing corals
df_status=read.csv("state_data.csv") #pulls in state data
df <- merge(df_surv, df_status, by.x = "ID", by.y = "Coral") #merge dataframes
#glimpse(df)
```

```{r Corellation Matrix}
#Select 
df_corr <- df %>% select(Mortality, Buried, Disease, Bleaching)
head(df_corr)
res <- cor(df_corr)
round(res,2)
```

```{r}
library(Hmisc)
res2 <-rcorr(as.matrix(df_corr))
res2
```
```{r correlation plot}
library(corrplot)
corrplot(res, type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45)
```
```{r}
# Insignificant correlation are crossed
corrplot(res2$r, type="upper", order="hclust", 
         p.mat = res2$P, sig.level = 0.01, insig = "blank")
# Insignificant correlations are leaved blank
corrplot(res2$r, type="upper", order="hclust", 
         p.mat = res2$P, sig.level = 0.01, insig = "blank")
```

```{r mortality by state}
#states column codes 
#Disease = 1
#Bleaching = 3
#Buried = 5
#Additive combos
#xtabs(~States+Mortality, data = df)
```
0 = no observed condition; 1 = diseased; 3 = bleaching; 5 = buried; and then additive combos to produce unique numbers. Condider recoding numbers to unique states.

```{r state by species}
xtabs(~Species+States, data = df)
```
Table shows the conditions by species where disease is a focus of this analysis including species specific patterns of disease susceptibility to SCTLD in particular.

```{r df long}
#Makes data.frame into long format for survival analyses
df_long <- gather(df,
                   key = "months",
                   value = "census",
                   X0:X21)
df1 <- df_long %>% mutate(months = as.numeric(gsub("X", "", months)))
```


```{r Cox Main Effects}
# Fit Cox Model
cox_fit<- coxph(Surv(months, census)~ Channel + Habitat + Dir, data=df1)
cox_fit
```

Double check, but exp(coef) is the relative hazards ratio and its se. Not sure what the HR is for habitat which should have 3 states?

```{r ANOVA Cox Main Effects}
Anova(cox_fit)
```
NS channel effect on mortality, but Habitat and Direction both significant.

```{r Habitat_survplot}
ggsurvplot(
  survfit(Surv(months, census)~ Habitat, data=df1),  # survfit object with calculated statistics
  pval = TRUE, pval.size = 4, # show p-value of log-rank test.
  #conf.int = TRUE, conf.int.style = "step", # customize style of confidence intervals
  xlab = "Time in Months",  # customize X axis label.
  break.time.by = 1,    # break X axis in time intervals by 200.
  ggtheme = theme_bw(), 
  xlim= c(12, 21),# customize plot and risk table with a theme.
  pval.coord = c(12,0.25),
  #surv.median.line = "hv",  # add the median survival pointer.
  legend = c("right"),
  legend.title = "Habitat",
  legend.labs = c("1", "2", "3"),    # change legend labels.
  #palette = c("#E7B800", "#2E9FDF"), # custom color palettes
  
  #risk.table = "percentage",  # absolute number and percentage at risk.
  #risk.table.fontsize = 3,# colour risk table text annotations.
  #risk.table.y.text = FALSE,# show bars instead of names in text annotations
  #ncensor.plot = TRUE      # plot the number of censored subjects at time t
)
```
```{r Dir_survplot}
ggsurvplot(
  survfit(Surv(months, census)~ Dir, data=df1),  # survfit object with calculated statistics
  pval = TRUE, pval.size = 4, # show p-value of log-rank test.
  #conf.int = TRUE, conf.int.style = "step", # customize style of confidence intervals
  xlab = "Time in Months",  # customize X axis label.
  break.time.by = 1,    # break X axis in time intervals by 200.
  ggtheme = theme_bw(), 
  xlim= c(12, 21),# customize plot and risk table with a theme.
  pval.coord = c(12,0.25),
  #surv.median.line = "hv",  # add the median survival pointer.
  legend = c("right"),
  legend.title = "Dir",
  legend.labs = c("North", "South"),    # change legend labels.
  #palette = c("#E7B800", "#2E9FDF"), # custom color palettes
  
  #risk.table = "absolute",  # absolute number and percentage at risk.
  #risk.table.fontsize = 3,# colour risk table text annotations.
  #risk.table.y.text = FALSE,# show bars instead of names in text annotations
  #ncensor.plot = TRUE      # plot the number of censored subjects at time t
)
```

Survival analyses focused on 9 most common species

```{r}
#15 or more
df_commons <- df1 %>% filter(Species %in% c("DCLI", "DSTO", "DSTR", "MCAV", "MMEA", "PAST", "SBOU", "SINT", "SSID")) #keeping only the 9 most common species 
#xtabs(~Species+Channel, data = df_commons)
```

```{r}
# Fit Cox Model
cox_fit2 <- coxph(Surv(months, census)~ Channel + Habitat + Dir + Species, data=df_commons)
cox_fit2
```

```{r}
Anova(cox_fit2)
```

Species then Habitat significant.


```{r Dir_survplot_sp}
ggsurvplot(
  survfit(Surv(months, census)~ Habitat, data=df_commons),  # survfit object with calculated statistics
  pval = TRUE, pval.size = 4, # show p-value of log-rank test.
  #conf.int = TRUE, conf.int.style = "step", # customize style of confidence intervals
  xlab = "Time in Months",  # customize X axis label.
  break.time.by = 1,    # break X axis in time intervals by 200.
  ggtheme = theme_bw(), 
  xlim= c(12, 21),# customize plot and risk table with a theme.
  pval.coord = c(12,0.25),
  surv.median.line = "hv",  # add the median survival pointer.
  legend = c("top"),
  legend.title = "Habitat",
  legend.labs = c("1", "2", "3"),    # change legend labels.
  #palette = c("black", "red"), # custom color palettes
  
  #risk.table = "percentage",  # absolute number and percentage at risk.
  #risk.table.fontsize = 3,# colour risk table text annotations.
  #risk.table.y.text = FALSE,# show bars instead of names in text annotations
  #ncensor.plot = TRUE      # plot the number of censored subjects at time t
)
```


```{r CommonSP_survplot}
ggsurvplot(
  survfit(Surv(months, census)~ Species, data=df_commons),  # survfit object with calculated statistics
  pval = TRUE, pval.size = 4, # show p-value of log-rank test.
  #conf.int = TRUE, conf.int.style = "step", # customize style of confidence intervals
  xlab = "Time in Months",  # customize X axis label.
  break.time.by = 1,    # break X axis in time intervals by 200.
  ggtheme = theme_bw(), 
  xlim= c(12, 21),# customize plot and risk table with a theme.
  pval.coord = c(12,0.25),
  surv.median.line = "hv",  # add the median survival pointer.
  legend = c("right"),
  legend.title = "Species",
  #legend.labs = c("Not Buried", "Buried"),    # change legend labels.
  #palette = c("black", "red"), # custom color palettes
  
  #risk.table = "percentage",  # absolute number and percentage at risk.
  #risk.table.fontsize = 3,# colour risk table text annotations.
  #risk.table.y.text = FALSE,# show bars instead of names in text annotations
  #ncensor.plot = TRUE      # plot the number of censored subjects at time t
)
```


```{r}
summary(survfit(Surv(months, census) ~ Species, data = df_commons), times = 21)
```
