---
title: "State Correlation Analyses"
output: github_document
always_allow_html: true
---

Markdown code for the survival analyses of the Port of Miami data examining the main effects of channel (i.e. dredging impact), reef habitat, and direction. Survival analyses used to tease apart species specific differences as well as the impact of state (or condition) specifically whether the corals were observed to be buried, bleached or diseased. Each variable is coded as 0 (absence) or 1 (presence) as well as in combination in the states column.

```{r setup}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE, error = FALSE, fig.path='Figures/', dev=c('png', 'pdf'))
```

All figures exported to Figures/folder as png and pdf

```{r libraries, include=FALSE}
library(knitr)
library(survival)
library(coda)
library(survminer)
library(dplyr)
library(ggplot2)
library(car)
library(tidyverse)
library(kableExtra)
library(ggpubr)
library(ggrepel)
library(patchwork)
#theme_bw()
```

```{r import data}
df=read.csv("9sp_corr.csv") #time by months; excludes missing corals
glimpse(df)
```

```{r bar plot all conditions facet}
df %>% filter(Condition %in% c("Bleaching","Disease","Buried", "Partial Burial","Partial Mortality","Mortality")) %>%
  mutate(prop = Value/Total.Individuals) %>%
  ggplot(aes(y=Species, x = prop)) +
  geom_bar(position = "stack", stat = "identity") +
  facet_wrap(~Condition)
```

```{r dot plot all conditions facet}
df %>% filter(Condition %in% c("Bleaching","Disease","Buried", "Partial Burial","Partial Mortality","Mortality")) %>%
  mutate(prop = Value/Total.Individuals*100) %>%
  ggdotchart(x="Species", y = "prop",
             color = "Condition",
             rotate = TRUE,
             add = "segments",
             dot.size = 6,
             label = "Value",
             repel = FALSE,
             xlab = "Top 9 Coral Species",
             ylab = "Percentage of Corals with Condition",
             font.label = list(color = "white", size = 9, vjust = 0.5),
             facet.by = "Condition")
```
```{r dot plot major conditions facet}
df %>% filter(Condition %in% c("Bleaching","Disease","Buried","Mortality")) %>%
  mutate(prop = Value/Total.Individuals*100) %>%
  ggdotchart(x="Species", y = "prop",
             color = "Condition",
             rotate = TRUE,
             add = "segments",
             dot.size = 6,
             label = "Value",
             repel = FALSE,
             xlab = "Top 9 Coral Species",
             ylab = "Percentage of Corals with Condition",
             font.label = list(color = "white", size = 9, vjust = 0.5),
             facet.by = "Condition")
```
```{r dot plot major conditions BEST}
df %>% filter(Condition %in% c("Bleaching","Disease","Buried","Mortality")) %>%
  mutate(prop = Value/Total.Individuals*100) %>%
  ggdotchart(x="Species", y = "prop",
             color = "Condition",
             rotate = TRUE,
             add = "segments",
             dot.size = 5,
             label = "Value",
             repel = FALSE,
             xlab = "Top 9 Coral Species",
             ylab = "Percentage of Corals with Condition",
             font.label = list(color = "white", size = 9, vjust = 0.5))
```

```{r}
library(reshape2)
df_w <- dcast(df, Species ~ Condition, value.var="Percentage")
kable(df_w)
```

```{r Corellation Matrix}
#Select 
df_corr <- df_w %>% select(Mortality, Buried, Disease, Bleaching, Normal, "Partial Mortality")
head(df_corr)
res <- cor(df_corr)
round(res,2)
```

```{r}
library(Hmisc)
res2 <-rcorr(as.matrix(df_corr))
res2
```

```{r correlation plot}
library(corrplot)
corrplot(res, type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45, 
         addCoef.col = "black",
         p.mat = res2$P, sig.level = 0.05,
         diag = FALSE)
```

```{r significant correlations BEST}
corrplot(res2$r, type="upper", order="hclust", 
         tl.col = "black", tl.srt = 45, 
         addCoef.col = "white",
         p.mat = res2$P, sig.level = 0.05, insig = "blank",
         diag = FALSE)
```

```{r relationship between disease and mortality}
ggscatter(df_w, y = "Mortality", x = "Disease",
       add = "reg.line",
       conf.int = TRUE,
       color = "Species", 
       add.params = list(color = "gray",
                         fill = "lightgray")
       ) + 
  stat_cor() #Add correlation coeff
```

