---
title: "Survival Analyses"
output: html_document
always_allow_html: true
---

Survival analyses of the Port of Miami data examining the main effects of channel (i.e. dredging impact), reef habitat, and direction for the full dataset followed by survival analyses of the top nine species dataset which includes species as a factor. 

Survival analyses used to tease apart species specific differences as well as the impact of state (or condition) specifically whether the corals were observed to be buried, bleached or diseased. Each variable is coded as 0 (absence) or 1 (presence) as well as in combination in the states column.

```{r setup}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE, error = FALSE)
```

All figures exported to Figures/folder as png and pdf

```{r libraries, include=FALSE}
library(knitr)
library(survival)
library(coda)
library(survminer)
library(dplyr)
library(ggplot2)
library(car)
library(tidyverse)
library(kableExtra)
library(ggpubr)
```

```{r import data}
df_surv=read.csv("survival_bymonth2.csv") #time by months; excludes missing corals
#df_surv<- df_surv %>% filter(Species %in% c("DCLI", "DSTO", "DSTR", "MCAV", "MMEA", "PAST", "SBOU", "SSID"))
#df_surv=subset(df_surv, Species!="MIS")
df_status=read.csv("coded_condition_data.csv") #pulls in state data
df <- merge(df_surv, df_status, by.x = "ID", by.y = "Coral") #merge dataframes

 #remove "SINT" which experience no mortality
#glimpse(df)
```

```{r df long}
#Makes data.frame into long format for survival analyses
df_long <- gather(df,
                   key = "months",
                   value = "census",
                   X0:X21)
df1 <- df_long %>% 
  mutate(months = as.numeric(gsub("X", "", months)))
```

```{r}
df1$Habitat <- as.character(df1$Habitat)
#Change Dir codes
df1$Direction <- recode_factor(df1$Direction, N = "North", S = "South")
#Change Habitat codes
df1$Habitat <- recode_factor(df1$Habitat, "1" = "HardBottom", "2" = "InnerReef", "3" = "OuterReef")
df1$Species <- recode_factor(df1$Species, "DCLI" = "PCLI")
df1$Species <- recode_factor(df1$Species, "DSTR" = "PSTR")
df_9sp <- df1 %>% mutate_if(is.character, factor)
```

```{r Cox Main Effects}
# Fit Cox Model
df_9sp$Dredge <- relevel(df_9sp$Dredge, ref ="control")
df_9sp$Species <- relevel(df_9sp$Species, ref ="SSID")
cox_fit_9sp<- coxph(Surv(months, census) ~ Dredge + Direction+ Habitat + Species, data=df_9sp)
cox_fit_9sp
```

```{r}
summary(cox_fit_9sp)
Anova(cox_fit_9sp)
```

```{r}
hr_plot_8sp<- ggforest(cox_fit_9sp, main = "Cox Proportional Hazard Model (8 species)")
hr_plot_8sp
ggsave("hr_plot_8sp.jpeg", height = 9, width=9)
```

```{r}
fit_sp_hab <- survfit(Surv(months, census) ~ Species + Habitat, data=df_9sp)
fit_sp_ch <- survfit(Surv(months, census) ~ Species + Dredge, data=df_9sp)
```

```{r}
ggsurv_main_onset <- ggsurvplot(fit_sp_hab,
           conf.int = TRUE,
           xlab = "Time in Months",
           break.time.by = 3,
           xlim = c(14, 21),
           ylim = c(0, 1),
          palette = get_palette("npg", 22))


surv_sp_hab <- ggsurv_main_onset$plot + theme_bw() + facet_grid(Habitat ~ Species) + theme(legend.position = "none")
surv_sp_hab
ggsave("surv_sp_hab.jpg")
```
```{r}
ggsurv_main_onset <- ggsurvplot(fit_sp_ch,
           conf.int = TRUE,
           xlab = "Time in Months",
           break.time.by = 3,
           xlim = c(14, 21),
           ylim = c(0, 1),
          palette = get_palette("npg", 22))


surv_sp_ch <- ggsurv_main_onset$plot + theme_bw() + facet_grid(Channel ~ Species) + theme(legend.position = "none")
surv_sp_ch
ggsave("surv_sp_ch.jpg")
```






```{r import data again}
df_surv2=read.csv("survival_bymonth.csv") #time by months; excludes missing corals
df_surv2<- df_surv2 %>% filter(Species %in% c("DCLI", "DSTO", "DSTR", "MCAV", "MMEA", "PAST", "SBOU", "SSID", "SINT"))
df_surv2=subset(df_surv2, Species!="MIS")
df_status2=read.csv("coded_condition_data.csv") #pulls in state data
df2 <- merge(df_surv2, df_status2, by.x = "ID", by.y = "Coral") #merge dataframes

 #remove "SINT" which experience no mortality
#glimpse(df)
```

```{r df long}
#Makes data.frame into long format for survival analyses
df_long2 <- gather(df2,
                   key = "months",
                   value = "census",
                   X0:X21)
df2 <- df_long2 %>% 
  mutate(months = as.numeric(gsub("X", "", months)))
```

```{r}
df2$Habitat <- as.character(df2$Habitat)
#Change Dir codes
df2$Direction <- recode_factor(df2$Direction, N = "North", S = "South")
#Change Habitat codes
df2$Habitat <- recode_factor(df2$Habitat, "1" = "HardBottom", "2" = "InnerReef", "3" = "OuterReef")
```

```{r}
#df2=subset(df1,Species==c("DCLI","DSTO","DSTR","MCAV","MMEA", "PAST", "SBOU", "SINT","SSID"))
  

res <- summary(survfit(Surv(months, census) ~ Species, data = df2), times = 21)
cols <- lapply(c(1:7, 8:11) , function(x) res[x])
tbl <- do.call(data.frame, cols)
#head(tbl)

tbl_surv <- tbl %>% dplyr::select(strata, n, surv, std.err)
tbl_surv

#tbl_surv%>% subset(strata==c("Species=MMEA", "Species=DSTR", "Species=DSTO","Species=DCLI", "Species=SBOU", "Species=MCAV", "Species=PAST", "Species=SSID", "Species=SINT"))

tbl_surv$strata  <- recode_factor(tbl_surv$strata , "Species=MMEA" = "MMEA", "Species=DSTR" = "PSTR", "Species=DSTO" = "DSTO","Species=DCLI" = "PCLI", "Species=SBOU" = "SBOU", "Species=MCAV" = "MCAV", "Species=PAST" = "PAST", "Species=SSID" = "SSID", "Species=SINT"="SINT")

tbl_surv$strata <- factor(tbl_surv$strata, levels = c("MMEA", "PSTR", "DSTO", "PCLI", "SBOU", "MCAV", "PAST", "SSID", "SINT"))


```

```{r species survival plot se}
# Black error bars - notice the mapping of 'group=supp' -- without it, the error
# bars won't be dodged!
ggplot(tbl_surv, aes(y = strata, x=surv, color = surv)) + 
  geom_errorbarh(aes(xmin=surv-std.err, xmax=surv+std.err), colour="black", width=.2, height=.2, size=.2) +
  geom_point(stat = "identity", size=2.5) + 
  theme_bw() +
  ylab("Coral Species") +
  xlab("Survival Probability (+-SE)") +
  scale_color_gradient(low = "red", high = "blue", name= "Survival 
Probability")
```


```{r species survival plot se}
ggsave("surv_sp_se.jpg",  width = 7, height = 5) 
    
```


