---
title: "State Correlation Analyses"
output: html_document
always_allow_html: true
---

Markdown code for the survival analyses of the Port of Miami data examining the main effects of channel (i.e. dredging impact), reef habitat, and direction. Survival analyses used to tease apart species specific differences as well as the impact of state (or condition) specifically whether the corals were observed to be buried, bleached or diseased. Each variable is coded as 0 (absence) or 1 (presence) as well as in combination in the states column.

```{r setup}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE, error = FALSE)
```


```{r libraries, include=FALSE}
#Load Packages
library(knitr)
library(survival)
library(coda)
library(survminer)
library(dplyr)
library(ggplot2)
library(car)
library(tidyverse)
library(kableExtra)
library(ggpubr)
library(ggrepel)
library(patchwork)
library("ggsci")
library("gridExtra")
library(rstatix)
theme_bw()
```

```{r import data, include=FALSE}
#Load Data
df=read.csv("9sp_corr.csv") #time by months; excludes missing corals
df %>% mutate_if(is.character, factor) 
df$Species <- factor(df$Species, levels = c("MMEA", "DSTR", "DSTO", "DCLI", "MCAV", "SSID", "SINT", "PAST", "SBOU"))

#Recode Factors
df$Condition<-recode_factor(df$Condition, "Normal"="No Conditions")
df$Condition<-recode_factor(df$Condition, "Buried"="Burial")
df$Species<-recode_factor(df$Species, "DCLI"="PCLI")
df$Species<-recode_factor(df$Species, "DSTR"="PSTR")

df$Condition <- factor(df$Condition, levels = c("Mortality", "Burial", "Bleaching", "Disease", "Partial Burial", "Partial Mortality", "No Conditions"))
```

Dot Plot
```{r dot plot all conditions facet}
df %>% filter(Condition %in% c("Bleaching","Disease","Burial", "Partial Burial","Partial Mortality","Mortality")) %>%
  mutate(prop = Value/Total.Individuals*100) %>%
  ggdotchart(x="Species", y = "prop",
             color = "Condition",
             rotate = TRUE,
             add = "segments",
             dot.size = 6,
             label = "Value",
             repel = FALSE,
             xlab = "Top 9 Coral Species",
             ylab = "Percentage of Corals with Condition",
             font.label = list(color = "white", size = 9, vjust = 0.5),
             facet.by = "Condition",
             palette = get_palette("npg", 6)
             )
```

Figure 7: Dot Plot
```{r dot plot major conditions facet}
Figure_7=df %>% filter(Condition %in% c("Bleaching","Disease","Burial","Mortality")) %>%
  mutate(prop = Value/Total.Individuals*100) %>%
  ggdotchart(x="Species", y = "prop",
             color = "Condition",
             rotate = TRUE,
             add = "segments",
             dot.size = 6,
             label = "Value",
             repel = FALSE,
             xlab = "Top 9 Coral Species",
             ylab = "% Corals with Condition (plus n corals in label)",
             font.label = list(color = "white", size = 9, vjust = 0.5),
             facet.by = "Condition",
             palette = get_palette("npg", 4))
Figure_7

#Save as JPG
ggsave("Figure_7.jpg", plot=Figure_7, width = 7, height = 5)
```

Cumulative Dot Plot by Species
```{r dot plot major conditions BEST}
sp_cond_dot <- df %>% filter(Condition %in% c("Bleaching","Disease","Buried","Mortality")) %>%
  mutate(prop = Value/Total.Individuals*100) %>%
  ggdotchart(x="Species", y = "prop",
             color = "Condition",
             rotate = TRUE,
             add = "segments",
             dot.size = 5,
             label = "Value",
             repel = FALSE,
             xlab = "Coral Species",
             sorting = c("none"),
             ylab = "% Corals with Condition\n (plus # of corals with condition in circle) ",
             font.label = list(color = "white", size = 9, vjust = 0.5),
             palette = get_palette("npg", 4),
             order = c("MMEA", "PSTR", "DSTO", "PCLI", "MCAV", "SSID", "SINT", "PAST", "SBOU")
             ) + theme_bw()
sp_cond_dot
ggsave("species_condition_dotchart.jpeg")
```

Set up for Correlation Plot
```{r}
library(reshape2)
df_w <- dcast(df, Species ~ Condition, value.var="Percentage")
df_w
```
Set up for Figure 8: Correlation Plot
```{r Corellation Matrix}
#Select 
df_corr <- df_w %>% dplyr::select(Mortality, Burial, Disease, Bleaching, "No Conditions", "Partial Mortality")
res <- cor(df_corr)
round(res,2)
```
Figure 8: Correlation Plot
```{r}
library(Hmisc)
cor.mat <- rcorr(as.matrix(df_corr))
cor.mat
cor3<- cor_plot(cor.mat,
  method = "number",
  type = "lower",
  significant.level = 0.05,
  insignificant = "cross"
)

cor3
ggsave("Figure_8.jpg", plot=cor3, width = 7, height = 5)
```

```{r}
#Select 
df_corr2 <- df_w %>% dplyr::select(Mortality, Burial, Disease, Bleaching)
res <- cor(df_corr2)
round(res,2)
```
Correlation Plot with only 4 conditions
```{r}
library(Hmisc)
cor.mat2 <- rcorr(as.matrix(df_corr2))
cor.mat2
cor2 <- cor_plot(cor.mat2,
  method = "number",
  type = "lower",
  significant.level = 0.05,
  insignificant = "cross"
)

cor2

ggsave("Corr_Plot.jpg", plot=cor2, width = 7, height = 5)
```


```{r}

```


```{r relationship between disease and mortality}
mxd_scatter <- ggscatter(df_w, y = "Mortality", x = "Disease", color = "Mortality",
       add = "reg.line",
       conf.int = TRUE,
       label = "Species",
       repel = TRUE,
       add.params = list(color = "grey",
                         fill = "lightgray")
       ) + stat_cor() + theme_bw() + gradient_color(c("blue", "red"))
mxd_scatter
ggsave("mortality_disease_scatter.jpg")
```














